Staged Abstract Interpreters

### 2019-09-11 paper/70_eval.tex
0-cfa

\vspace{-5pt}
\paragraph{Implementation and Evaluation Environment} We implement the abstract
interpreters for control-flow analysis, i.e., a value-flow analysis that
computes which $\lambda$-terms can be called at each call-site.
With other suitable abstract domains, the abstract interpreter
can also be extended to other analyses.
A front-end desugars Scheme to a small core language that makes the analyzer
easy to implement. We implement two common monovariant 0-CFA-like analyses: one
is equipped with store-widening (0CFA-SW for short), the other one is not (0CFA
for short). It is traditionally expected that the store-widened analysis be faster than the
alternative.  Our prototype implementation currently generates Scala code. The
generated Scala code will be compiled and executed on the JVM, though it is possible to
generate C/C++ code in the future. In the experiments, we implement several
optimizations mentioned in Section \ref{staged_ds}, specifically, including the
selective caching scheme and IR-level rewriting rules (for pairs, lists, sets,
and maps) that exploit partially-static data. These optimizations are useful to
reduce the size of generated code and improve performance.

@@

Figure~\ref{evaluation_result} shows the evaluation result, comparing the
performance improved by staging on the two monovariant CFAs. The
\textit{unstaged} and \textit{staged} columns show the median time (excluding
pre-compilation) to finish the analysis. The column
$\frac{\text{unstaged}}{\text{staged}}$ shows the improvement. The dash @-@ in
some cells indicates timeout.  As expected, the staged version significantly
outperforms the unstaged version on all benchmarks.
%We observe that simple value domains used in the benchmark usually leads to
%better speed-ups, for example, the @kcfa-worst@ series only uses lambdas and
%booleans.
For 0CFA without store-widening, we observe an average speedup of ~17 times;
for 0CFA with store-widening, the average speedup is ~7 times.
\texttt{kcfa-worst-64} has the highest speedup among all the tests. Although it
has a comparably large size of AST, it has a rather simple structure which only
involves booleans, $\lambda$-terms, and applications. We conjecture that the simple
domain and structure allow the generated code to be efficiently optimized by the
JVM. However, on the corresponding 0CFA-SW, we do not observe such a significant
speedup. Due to the complexity of generated code and next-stage running platform
(JVM), we leave the detailed empirical analysis of which static analysis can
benefit more from the specialization as future work.

>>>>

\vspace{-5pt}
\paragraph{Implementation and Evaluation Environment} We implement the abstract
interpreters for control-flow analysis, i.e., a value-flow analysis that
computes which $\lambda$-terms can be called at each call-site.
With other suitable abstract domains, the abstract interpreter
can also be extended to other analyses.
A front-end desugars Scheme to a small core language that makes the analyzer
easy to implement. We implement two common monovariant 0-CFA-like analyses: one
is equipped with store-widening (0-CFA-SW for short), the other one is not (0-CFA
for short). It is traditionally expected that the store-widened analysis be faster than the
alternative.  Our prototype implementation currently generates Scala code. The
generated Scala code will be compiled and executed on the JVM, though it is possible to
generate C/C++ code in the future. In the experiments, we implement several
optimizations mentioned in Section \ref{staged_ds}, specifically, including the
selective caching scheme and IR-level rewriting rules (for pairs, lists, sets,
and maps) that exploit partially-static data. These optimizations are useful to
reduce the size of generated code and improve performance.

@@

Figure~\ref{evaluation_result} shows the evaluation result, comparing the
performance improved by staging on the two monovariant CFAs. The
\textit{unstaged} and \textit{staged} columns show the median time (excluding
pre-compilation) to finish the analysis. The column
$\frac{\text{unstaged}}{\text{staged}}$ shows the improvement. The dash @-@ in
some cells indicates timeout.  As expected, the staged version significantly
outperforms the unstaged version on all benchmarks.
%We observe that simple value domains used in the benchmark usually leads to
%better speed-ups, for example, the @kcfa-worst@ series only uses lambdas and
%booleans.
For 0-CFA without store-widening, we observe an average speedup of ~17 times;
for 0-CFA with store-widening, the average speedup is ~7 times.
\texttt{kcfa-worst-64} has the highest speedup among all the tests. Although it
has a comparably large size of AST, it has a rather simple structure which only
involves booleans, $\lambda$-terms, and applications. We conjecture that the simple
domain and structure allow the generated code to be efficiently optimized by the
JVM. However, on the corresponding 0-CFA-SW, we do not observe such a significant
speedup. Due to the complexity of generated code and next-stage running platform
(JVM), we leave the detailed empirical analysis of which static analysis can
benefit more from the specialization as future work.
